SPARQL on the Open, Decentralised Web

### 2016-08-03 dissertation/cache-proxy.tex
Some fixes closes #17

It is further interesting to note the contrast to the HERMES
\cite{adali1996query} system's concept of invariants discussed in
Section~\ref{sec:relcache}: a SPARQL query cache be oblivious
to domain knowledge in a CDN deployment and therefore shouldn't assume any invariants,
except in the possible but unlikely case where the remote server
declares an infinite freshness lifetime for the result of a certain
graph pattern. A forward proxy, on the other hand, may have a deeper
understanding on the needs of the client, and could possibly implement
invariants. The Attean framework accommodates for this.

Figure~\ref{fig:messaging} illustrates the case where HTTP messages
are being passed, where the cache is assumed to reside on a
intermediate caching proxy in the Internet, i.e. a CDN. The figure
may be viewed as having time on the $x$-axis, with the arrows pointing
down being requests and the arrows point up being responses. In this
example, the client first makes a request (request \circled{1}), which
is passed through the proxy because it has not yet cached any
responses. The response (response \circled{1}) is then also
transmitted directly from the SPARQL Endpoint at the database back to
the client. The entire response may be cached on the proxy in case it
can be reused in its entirety. However, the main point is that the
request will be analysed by the proxy in parallel to being sent to the
server, and another request \circled{2} for a triple pattern fragment
that the analyser thinks will enable the proxy to assist the server
the best for future requests, is sent. This analysis is discussed in
Section~\ref{sec:analpre}. The response \circled{2} is not sent back
to the client, but will enter the cache. Assuming that the analyser
was successful and that the cache can be used to answer request
\circled{3}. In that case, the proxy will send rewritten queries for
the rest of the data it needs to be able to answer the query to the
origin server, indicated by the dashed arrows. This may be zero or more SPARQL
query, TPF requests or a combination thereof, depending on what the
planner finds least expensive. The proxy will then evaluate the entire
query and send the final result back as response \circled{3}. In
addition, the analyser may have identified another Triple Pattern
Fragment that it takes as likely to be useful for future requests, and
sends a request and caches the response \circled{4}. At some point, it
may be able to answer the entire query using only cached results, as
illustrated by request and response \circled{5}.

@@

As can be imagined from the fact that a BGP with just to triple
patterns result in 4 plan trees, a naive query planner may generate a
large number of plans for complex queries. In addition, the proposed
system will generate even more plans, since it considers three sources
of the same data, the cache, the remote endpoint and an LDF
server. Various methods have been developed in the literature to limit
the growth of plans. In addition to a naive planner, we also have an
Iterative Dynamic Programming planner
\cite{Kossmann:2000:IDP:352958.352982}, which prunes plans
aggressively based on their estimated cost in each iteration of query
plan generation.

@@

Also of importance are the APIs to compose stores. Stores in Attean
can be triplestores or quadstores, they can be mutable,
bulkupdateable, can enable the use of ETags or modification times for
conditional requests (as defined in RFC7232 \cite{rfc7234}), and
enables further extensions. Each of these primitives is represented
by a role that implements default functionality or simply requires
such functionality to be required. For instance, a quadstore is
required by \pmodule{Attean::API::QuadStore} to implement the
\pcode{get\_quads} method, which can take an RDF term as one or more
of the arguments and variables for the rest, and should return an
iterator over the quads that were matched. Based on an implementation
of the \pcode{get\_quads} method, the \pmodule{Attean::API::QuadStore}
role provides default implementations of methods \pcode{count\_quads},
\pcode{count\_quads\_estimate}, \pcode{get\_graphs} and
\pcode{size}. These methods may be overridden with more efficient
implementations when composing a class. Composing other roles will
require further methods to be implemented.

@@

It is clear that the vast majority of time spent is due to my cache
implementation. When using the Redis cache, I serialise RDF terms as
N-Triples, and when they are retrieved, they have to be parsed and
instantiated as objects for the query engine. Remarkably, parsing of
Internationalized Resource Identifiers (IRIs) is the single most
time-consuming task. While each parse does not take much time, it
happens more than 13~million times. The reason why parsing IRIs
happens is that a certain amount of parsing is required to compare
them \cite{rfc3987}, and the \pmodule{IRI} module is a rigorous
implementation of the standard that does syntax based resolution if
the if an object is constructed with a base. IRI parsing happens not
only when an \pmodule{Attean::IRI} object is constructed, but may also
happen when an \pmodule{Attean::Literal} object is constructed, since
the data type is a IRI.

>>>>

It is further interesting to note the contrast to the HERMES
\cite{adali1996query} system's concept of invariants discussed in
Section~\ref{sec:relcache}: a SPARQL query cache should be oblivious
to domain knowledge in a CDN deployment and therefore shouldn't assume any invariants,
except in the possible but unlikely case where the remote server
declares an infinite freshness lifetime for the result of a certain
graph pattern. A forward proxy, on the other hand, may have a deeper
understanding of the needs of the client, and could possibly implement
invariants. The Attean framework accommodates for this.

Figure~\ref{fig:messaging} illustrates the case where HTTP messages
are being passed, where the cache is assumed to reside on a
intermediate caching proxy in the Internet, i.e. a CDN. The figure
may be viewed as having time on the $x$-axis, with the arrows pointing
down being requests and the arrows point up being responses. In this
example, the client first makes a request (request \circled{1}), which
is passed through the proxy because it has not yet cached any
responses. The response (response \circled{1}) is then also
transmitted directly from the SPARQL Endpoint at the database back to
the client. The entire response may be cached on the proxy in case it
can be reused in its entirety. However, the main point is that the
request will be analysed by the proxy in parallel to being sent to the
server, and another request \circled{2} for a triple pattern fragment
that the analyser thinks will enable the proxy to assist the server
the best for future requests, is sent. This analysis is discussed in
Section~\ref{sec:analpre}. The response \circled{2} is not sent back
to the client, but will enter the cache. We then assume that the analyser
was successful and that the cache can be used to answer request
\circled{3}. In that case, the proxy will send rewritten queries for
the rest of the data it needs to be able to answer the query to the
origin server, indicated by the dashed arrows. This may be zero or more SPARQL
queries, TPF requests or a combination thereof, depending on what the
planner finds least expensive. The proxy will then evaluate the entire
query and send the final result back as response \circled{3}. In
addition, the analyser may have identified another Triple Pattern
Fragment that it takes as likely to be useful for future requests, and
sends a request and caches the response \circled{4}. At some point, it
may be able to answer the entire query using only cached results, as
illustrated by request and response \circled{5}.

@@

As can be imagined from the fact that a BGP with just two triple
patterns result in 4 plan trees, a naive query planner may generate a
large number of plans for complex queries. In addition, the proposed
system will generate even more plans, since it considers three sources
of the same data, the cache, the remote endpoint and an LDF
server. Various methods have been developed in the literature to limit
the growth of plans. In addition to a naive planner, we also have an
Iterative Dynamic Programming planner
\cite{Kossmann:2000:IDP:352958.352982}, which prunes plans
aggressively based on their estimated cost in each iteration of query
plan generation.

@@

Also of importance are the APIs to compose stores. Stores in Attean
can be triplestores or quadstores, they can be mutable,
bulkupdateable, can enable the use of ETags or modification times for
conditional requests (as defined in RFC7232 \cite{rfc7234}), and
enables further extensions. Each of these primitives is represented
by a role that implements default functionality or simply requires
such functionality to be present. For instance, a quadstore is
required by \pmodule{Attean::API::QuadStore} to implement the
\pcode{get\_quads} method, which can take an RDF term as one or more
of the arguments and variables for the rest, and should return an
iterator over the quads that were matched. Based on an implementation
of the \pcode{get\_quads} method, the \pmodule{Attean::API::QuadStore}
role provides default implementations of methods \pcode{count\_quads},
\pcode{count\_quads\_estimate}, \pcode{get\_graphs} and
\pcode{size}. These methods may be overridden with more efficient
implementations when composing a class. Composing other roles will
require further methods to be implemented.

@@

It is clear that the vast majority of time spent is due to my cache
implementation. When using the Redis cache, I serialise RDF terms as
N-Triples, and when they are retrieved, they have to be parsed and
instantiated as objects for the query engine. Remarkably, parsing of
Internationalized Resource Identifiers (IRIs) is the single most
time-consuming task. While each parse does not take much time, it
happens more than 13~million times. The reason why parsing IRIs
happens is that a certain amount of parsing is required to compare
them \cite{rfc3987}, and the \pmodule{IRI} module is a rigorous
implementation of the standard that does syntax based resolution
if an object is constructed with a base. IRI parsing happens not
only when an \pmodule{Attean::IRI} object is constructed, but may also
happen when an \pmodule{Attean::Literal} object is constructed, since
the data type is a IRI.

### 2016-08-03 dissertation/introdefhistory.tex
Frame the big problem statement, closes #7
Go through text flow #18
Some fixes closes #17

As noted in Section~\ref{sec:contribsum}, the scope of the study is
very broad, ranging from the practical programming details to
philosophy of science. With a scope of this breadth, the investigation
is shallow, it is hoped that some key insights will provoke further
investigation rather than being the end of any conversations.

The introduction provided a very basic view of the technology, this
chapter will further discuss some important aspects of the Semantic
Web, as it provides guidance for the investigation detailed in the
papers. The overarching problem this work tries to address is this:

\begin{quote}
To achieve a Semantic Web where anyone can say anything about
anything, and where everyone is enabled to analyse data that have been
contributed, there is a difficult balance between centralising
infrastructure and decentralising it, and between processing on
clients and servers.
\end{quote}

As such, the following serves to tie the published papers into a
whole, provide an overarching rationale and problem statements,
overview of the contributions and a discussion of the problematic
sides of the study seen from a philosophy of science viewpoint. It is
structured as follows: Section~\ref{sec:nodef} discusses the
possibility of defining the Semantic Web,
Section~\ref{sec:objectivity} discusses the problem of maintaining
objectivity, Section~\ref{sec:history} attempts to satisfy the
previous discussion by detailing the author's personal involvement in
the Semantic Web. Section~\ref{sec:motivation} has an overview of the
motivation behind this work.  Section~\ref{sec:prelims} provides some
further technical details and Section~\ref{sec:related} puts the
present work into the context within the existing
literature. Section~\ref{sec:problems} summarises important research
problems and details the motivation.
Sections~\ref{sec:papersincontext}~and~\ref{sec:unpublished} details
how the contributions address these problems and discusses weaknesses
of the papers with the benefit of hindsight. Future work is discussed
in Section~\ref{sec:future}. Key insights are finally summarised in
Section~\ref{sec:conclusions}.

@@

In 1996, I started the Web pages of the \textit{Norwegian Skeptics Society}
\texttt{skepsis.no}, and therefore got the control over an entire Web
server. That was another revelation, as I could actually run code, and
from that arose the need to not just learn by ``View Source'', but
start to read the specifications, notably HTML, CSS and
HTTP. Appreciating the design of these specifications took a while,
but eventually, the value of orthogonal specifications and separation
of concerns became apparent. Also, at the same time, I realised that I
could be standing on the shoulders of giants by using Free Software.

>>>>

As noted in Section~\ref{sec:contribsum}, the scope of the study is
very broad, ranging from the practical programming details to
philosophy of science. With a scope of this breadth, the investigation
is shallow, it is hoped that some key insights will provoke further
investigation rather than being the end of any conversations. 
The overarching problem this work tries to address is this:

\begin{framed}
\begin{quote}
To achieve a Semantic Web where anyone can say anything about
anything, and where everyone is enabled to analyse data that have been
contributed, there is a difficult balance between centralising
infrastructure and decentralising it, and between processing on
clients and servers.
\end{quote}
\end{framed}

The introduction provided a very basic view of the technology, this
chapter will further discuss some important aspects of the Semantic
Web, as it provides guidance for the investigation detailed in the
papers.

As such, the following serves to tie the published papers into a
whole, provide an overarching rationale and problem statements,
overview of the contributions and a discussion of the problematic
sides of the study seen from a philosophy of science viewpoint. It is
structured as follows: Section~\ref{sec:nodef} discusses the
possibility of defining the Semantic Web,
Section~\ref{sec:objectivity} discusses the problem of maintaining
objectivity, Section~\ref{sec:history} attempts to satisfy the
previous discussion by detailing the author's personal involvement in
the Semantic Web. Section~\ref{sec:motivation} has an overview of the
motivation behind this work.  Section~\ref{sec:prelims} provides some
further technical details and Section~\ref{sec:related} puts the
present work into the context within the existing
literature. Section~\ref{sec:problems} summarises important research
problems and details the motivation.
Sections~\ref{sec:papersincontext}~and~\ref{sec:unpublished} detail
how the contributions address these problems and discusses weaknesses
of the papers with the benefit of hindsight. Future work is discussed
in Section~\ref{sec:future}. Key insights are finally summarised in
Section~\ref{sec:conclusions}.

@@

In 1996, I started the Web pages of the \textit{Norwegian Skeptics
  Society} using the domain
\texttt{skepsis.no}, and therefore got the control over an entire Web
server. That was another revelation, as I could actually run code, and
from that arose the need to not just learn by ``View Source'', but
start to read the specifications, notably HTML, CSS and
HTTP. Appreciating the design of these specifications took a while,
but eventually, the value of orthogonal specifications and separation
of concerns became apparent. Also, at the same time, I realised that I
could be standing on the shoulders of giants by using Free Software.
